THESIS ANALYSES

```{r}
require("vegan")
require("dplyr")
require("nlme")
require("reshape2")
require("ecodist")
require("ggplot2")
require("ade4")
require("png")
require("MASS")
```

```{r}
#load design file without source tanks

design <- read.csv("C:/Users/astuc/Desktop/GitHub/CSI_Dispersal/data/CSI_Design_ENV_NoSourceTanks.csv",row.names=1)
design

#load design file with microbes 
design_crobes <- read.csv("C:/Users/astuc/Desktop/GitHub/CSI_Dispersal/data/design_CSI.csv",row.names=1)

#subset without mock community
design_crobes <- design_crobes[-c(grep("mock community", design_crobes$CSI_ID)), ]
dim(design_crobes)
```

 set treatments
 
```{r}
sal_trt <- design$Salinity
levels(sal_trt) <- c("0","5","9","13")
disp_trt <- design$Dispersal
levels(disp_trt) <- c("2","3")
```

```{r}
#set source mothur tools
source("C:/Users/astuc/Desktop/GitHub/CSI_Dispersal/bin/DiversityFunctions.R")
source("C:/Users/astuc/Desktop/GitHub/CSI_Dispersal/bin/MothurTools.R")

csi_otu <- read.otu("C:/Users/astuc/Desktop/GitHub/CSI_Dispersal/data/CSI.shared")
dim(csi_otu)
```

remove extra site in design (CSI041)

```{r}
missing <- setdiff(rownames(design_crobes), rownames(csi_otu))
design_crobes <- design_crobes[-(which(rownames(design_crobes) == missing)), ]
dim(design_crobes)

design <- design[-(which(rownames(design) == missing)), ]
dim(design)
```

id source tanks where number = 1,2,3

```{r}
temp <- rownames(design_crobes[which(design_crobes$Number %in% c(1,2,3)), ])
```

remove samples from tank numbers 1,2,3

```{r}
bac.design <- design_crobes[-(which(rownames(design_crobes) %in% temp)), ]
design_crobes2 <- droplevels(bac.design)
dim(design_crobes2)
```

remove source tanks where number = 1,2,3

```{r}
csi_otu_table <- csi_otu[-(which(rownames(csi_otu) %in% temp)), ]
dim(csi_otu_table)
```

remove otu's w/ < 2 occurrences across all sites

```{r}
otu_removal <- csi_otu_table[, which(colSums(csi_otu_table) >= 2)]
dim(otu_removal)
```

rarefy abundances

```{r}
rowSums(otu_removal)
otu_rarefy <- rrarefy(otu_removal, 13000)
```

removed low abundance samples

```{r}
csi_low_remov <- otu_removal[which(rowSums(otu_removal) >= 13000), ]
dim(csi_low_remov)
```

odd sites in bacterial composition data (101) and remove 041 in design file

```{r}
odd.sites <- c("CSI101")

otu_final <- csi_low_remov[setdiff(rownames(csi_low_remov), odd.sites), ]
design_final <- design[setdiff(rownames(design), odd.sites), ]

all.equal(rownames(design_final), rownames(otu_final))
```

create presence absence matrix

```{r}
csi_pres_abs <- (otu_final > 0) * 1
```

create relative abundance matrices

```{r}
csi_relabun <- otu_final
for (i in 1:dim(otu_final)[1]) {
csi_relabun[i, ] <- otu_final[i, ]/sum(otu_final[i,
])
}
```

log transform rel abun's

```{r}
csi_log <- decostand(csi_relabun, method = "log")
```

import tax file (note: this won't work? 8/15)
Error received when running line 138 code read.tax: Error in colsplit(tax_raw[, 3], split = "\\;", names = c("Domain", "Phylum", : unused argument (split = "\\;")

```{r}
csi_tax <- read.tax(taxonomy="C:/Users/astuc/Desktop/GitHub/CSI_Dispersal/data/CSI.0.03.cons.taxonomy")

write.table(csi_tax, file = "tax.csv", sep ",",
            col.names = NA)
```

bind design and bact files

```{r}
new_csi <- cbind(design_final,csi_relabun)
df <- decostand(csi_relabun, method="hellinger")
```

drop missing NA for salinity

```{r}
csi_dat <- na.omit(new_csi)

df_bact <- csi_dat[,-c(1:15)]
df_env <- csi_dat[,c(1:15)]
df_nutr <- csi_dat[,c(9:11)]
```

make OTU relative?

```{r}
csi_fin <- df_bact[,colSums(df_bact) >0]
```

not sure about this?

```{r}
csi_fin_rel <- csi_fin
	for (i in 1:26){
		csi_fin_rel[i,]<-csi_fin[i,]/sum(csi_fin[i,])
	}
rowSums(csi_fin_rel)
```

```{r}
bact_dist <- vegdist(csi_fin_rel)
env_dist <- vegdist(scale(df_env), "euclid") #did not work
nutr_dist <- vegdist(df_nutr)
```

alpha diversity (shannon weiner)

```{r}
a_div <- diversity(csi_fin)
a_div
```

beta diversity - confusing

"Function betadiver finds all indices reviewed by Koleff et al. (2003). All these indices could be found with function designdist which uses different notation, but the current function provides a conventional shortcut. The function only finds the indices. The proper analysis must be done with functions such as betadisper, adonis or mantel.""

```{r}
b_div <- betadiver(csi_fin, method = "w")
b_div
```




