THESIS ANALYSES

```{r}
require("vegan")
require("dplyr")
require("nlme")
require("reshape2")
require("ecodist")
require("ggplot2")
require("ade4")
require("png")
require("MASS")
require("grid")
require("ape")
require("png")
require("picante")

myColors <- c("#448844", "#33CC33", "#CCFF00", "#FFF000", "#FF9933", "#A9A9A9")
names(myColors) <- c("BalledBurlapped", "Bareroot", "Seedling",
"Acorn", "Seedbank", "Reference")
```

```{r}
#load design file without source tanks

se <- function(x, ...) {
sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))
}
ci <- function(x, ...) {
1.96 * sd(x, na.rm = TRUE)
}

design <- read.csv("C:/Users/astuc/Desktop/GitHub/CSI_Dispersal/data/CSI_Design_ENV_NoSourceTanks.csv",row.names=1)
design

#load design file with microbes 
design_crobes <- read.csv("C:/Users/astuc/Desktop/GitHub/CSI_Dispersal/data/design_CSI.csv",row.names=1)

#subset without mock community
design_crobes <- design_crobes[-c(grep("mock community", design_crobes$CSI_ID)), ]
dim(design_crobes)
```

 set treatments
 
```{r}
sal_trt <- design$Salinity
levels(sal_trt) <- c("0","5","9","13")
disp_trt <- design$Dispersal
levels(disp_trt) <- c("2","3")
```

```{r}
#set source mothur tools
source("C:/Users/astuc/Desktop/GitHub/CSI_Dispersal/bin/DiversityFunctions.R")
source("C:/Users/astuc/Desktop/GitHub/CSI_Dispersal/bin/MothurTools.R")

csi_otu <- read.otu("C:/Users/astuc/Desktop/GitHub/CSI_Dispersal/data/CSI.shared")
dim(csi_otu)
```

remove extra site in design (CSI041)

```{r}
missing <- setdiff(rownames(design_crobes), rownames(csi_otu))
design_crobes <- design_crobes[-(which(rownames(design_crobes) == missing)), ]
dim(design_crobes)

design <- design[-(which(rownames(design) == missing)), ]
dim(design)
```

id source tanks where number = 1,2,3

```{r}
temp <- rownames(design_crobes[which(design_crobes$Number %in% c(1,2,3)), ])
```

remove samples from tank numbers 1,2,3

```{r}
bac.design <- design_crobes[-(which(rownames(design_crobes) %in% temp)), ]
design_crobes2 <- droplevels(bac.design)
dim(design_crobes2)
```

remove source tanks where number = 1,2,3

```{r}
csi_otu_table <- csi_otu[-(which(rownames(csi_otu) %in% temp)), ]
dim(csi_otu_table)
```

remove otu's w/ < 2 occurrences across all sites

```{r}
otu_removal <- csi_otu_table[, which(colSums(csi_otu_table) >= 2)]
dim(otu_removal)
```

rarefy abundances

```{r}
rowSums(otu_removal)
otu_rarefy <- rrarefy(otu_removal, 13000)
```

removed low abundance samples

```{r}
csi_low_remov <- otu_removal[which(rowSums(otu_removal) >= 13000), ]
dim(csi_low_remov)
```

odd sites in bacterial composition data (101) and remove 041 in design file

```{r}
odd.sites <- c("CSI101")

otu_final <- csi_low_remov[setdiff(rownames(csi_low_remov), odd.sites), ]
design_final <- design[setdiff(rownames(design), odd.sites), ]

all.equal(rownames(design_final), rownames(otu_final))
```

create presence absence matrix

```{r}
csi_pres_abs <- (otu_final > 0) * 1
```

create relative abundance matrices

```{r}
csi_relabun <- otu_final
for (i in 1:dim(otu_final)[1]) {
csi_relabun[i, ] <- otu_final[i, ]/sum(otu_final[i,
])
}
```

log transform rel abun's

```{r}
csi_log <- decostand(csi_relabun, method = "log")
```

import tax file (note: this won't work? 8/15)
Error received when running line 138 code read.tax: Error in colsplit(tax_raw[, 3], split = "\\;", names = c("Domain", "Phylum", : unused argument (split = "\\;")

```{r}
csi_tax <- read.tax(taxonomy="C:/Users/astuc/Desktop/GitHub/CSI_Dispersal/data/CSI.0.03.cons.taxonomy")
```

bind design and bact files

```{r}
new_csi <- cbind(design_final,csi_relabun)
df <- decostand(csi_relabun, method="hellinger")
```

drop missing NA for salinity

```{r}
csi_dat <- na.omit(new_csi)

df_bact <- csi_dat[,-c(1:15)]
df_env <- csi_dat[,c(1:15)]
df_nutr <- csi_dat[,c(9:11)]
```

make OTU relative?

```{r}
csi_fin <- df_bact[,colSums(df_bact) >0]
```

not sure about this?

```{r}
csi_fin_rel <- csi_fin
	for (i in 1:26){
		csi_fin_rel[i,]<-csi_fin[i,]/sum(csi_fin[i,])
	}
rowSums(csi_fin_rel)
```

```{r}
bact_dist <- vegdist(csi_fin_rel)
env_dist <- vegdist(scale(df_env), "euclid") #did not work
nutr_dist <- vegdist(df_nutr)
```

alpha diversity (shannon weiner)

```{r}
a_div <- diversity(csi_fin)
a_div
```

beta diversity - confusing

"Function betadiver finds all indices reviewed by Koleff et al. (2003). All these indices could be found with function designdist which uses different notation, but the current function provides a conventional shortcut. The function only finds the indices. The proper analysis must be done with functions such as betadisper, adonis or mantel.""

```{r}
b_div <- betadiver(csi_fin, method = "w")
b_div
```

attempting diversity metrics (from stormwater) 
THINK ABOUT LOOKING AT DIV ACCORDING TO BOTH SALINITY AND DISPERSAL
HOW TO DO THIS?

```{r}
fisher <- fisher.alpha(otu_rarefy)
fisher

richness <- rowSums((otu_rarefy >= 1))
richness

shannon <- diversity(otu_rarefy, "shannon")
shannon

simp.even <- apply(otu_rarefy, 1, simp_even)
simp.even

colnames(design.r)

```

attempting diversity function plots (from stormwater)

```{r}
design_final
otu_final

otu_n_design <- cbind(design_final,otu_final)

design.r <- design_final[which(rownames(design_final) %in% rownames(otu_final)), ]

design.r

design.disp <- design.r[design.r$Dispersal != "W", ]
design.sal <- design.r[design.r$Salinity != "w", ]
design.date <- design.r[design.r$Date2 != "W", ]

disp.bac <- as.data.frame(matrix(NA, dim(design.disp)[1], 10))
sal.bac <- as.data.frame(matrix(NA, dim(design.sal)[1], 10))
date.bac <- as.data.frame(matrix(NA, dim(design.date)[1], 10))

colnames(disp.bac) <- c("Field_ID","Replicate","Date2","Treatment","Salinity","Dispersal","S","F_Alpha","S_Even","Shannon")

disp.bac$Sample <- rownames(design.disp)
disp.bac$Field_ID <- design.disp$Field_ID
disp.bac$Replicate <- design.disp$Replicate
disp.bac$Date2 <- design.disp$Date2
disp.bac$Treatment <- design.disp$Treatment
disp.bac$Salinity <- design.disp$Salinity
for(i in 1:length(rownames(design.disp))){
  temp <- rownames(design.disp)[i]
  disp.bac$S[i] <- richness[which(names(richness) == temp)]
  disp.bac$F_Alpha[i] <- fisher[which(names(fisher) == temp)]
  disp.bac$S_Even[i] <- simp.even[which(names(simp.even) == temp)]
  disp.bac$Shannon[i] <- shannon[which(names(shannon) == temp)]
}

disp.bac.m <- melt(disp.bac, id = c("Sample","Date2","Replicate","Salinity"))
disp.bac.c <- cast(data = disp.bac.m, Sample + Date2 ~ variable, c(mean,se), na.rm=T)

disp.bac.c <- as.data.frame(disp.bac.c)
disp.bac.m <- as.data.frame(disp.bac.m)

png(filename="../analyses/TEST_Div_Fxn.png",
    width = 1200, height = 800, res = 96*2)

test_plot <- barplot(disp.bac.c[, 5],
                     ylab="",xlab="",
                     lwd=3,col="gray",cex.lab=1.25,cex.names = 1.25, 
                     ylim = c(0,4500),yaxt="n",
                     space=c(1,0.25,1,0.25,1,0.25),
                     density=c(-1,15,-1,15,-1,15))
arrows(x0=test_plot, y0=disp.bac.c[, 5], y1 = disp.bac.c[, 5] - disp.bac.c[, 6], angle=90, length=0.1, lwd=2)
arrows(x0=test_plot, y0=disp.bac.c[, 5], y1 = disp.bac.c[, 5] + disp.bac.c[, 6], angle=90, length=0.1, lwd=2)
axis(side=2, labels=T,lwd.ticks=2,las=2,lwd=2,at=c(seq(0,150,75)))
abline(h=0,lwd=2,lty=3)

#need to figure this out -- how to make diversity figures? very different from stormwater example... maybe need another method
```

CHANGING PACE - RELATIONSHIP B/T COMM COMP AND DECOMP RATES
dbRDA and Mantel

Mantel - similarity between crobe and zoop communities? (based off mantel test in IL Wetlands looking at crobe/plant)

```{r}
zoop_csi <- read.csv("../data/zoop_CSI.csv")

all.equal(rownames(zoop_csi),rownames(csi_relabun))

zoop_csi
#which relabun matrix to use? may have to re-create
```

